---
title: "Pentanucleotide change sigs PCAWG"
author: "Hilmar Berger"
date: "3/21/2020"
output: html_document
---

```{r, message=FALSE}
library(data.table)
library(reshape2)
library(ggplot2)
library(pheatmap)
```

# Data import

## PCAWG WGS pentanucleotide changes

```{r}
# Pentanucleotide change counts for each sample
pnc_data = fread("../../../Data/External/PCAWG/WGS_PCAWG.1536.csv")
sids = colnames(pnc_data)[3:ncol(pnc_data)]
tmp = strsplit(sids, "::")
ed = data.frame(row.names=sids, cancer_entity = unlist(sapply(tmp, `[`,1)), specimenID=unlist(sapply(tmp, `[`,2)), stringsAsFactors = F)

pnc_mat = t(as.matrix(pnc_data[,3:ncol(pnc_data)]))
colnames(pnc_mat) = paste0(substr(pnc_data$Pentanucleotide,1,2),"[",pnc_data$`Mutation type`,"]", substr(pnc_data$Pentanucleotide,4,5))

pnc_mat_rel = sweep(pnc_mat, 1, apply(pnc_mat,1,sum),"/")
```

## PCAWG WGS trinucleotide based signature contributions

```{r}
sbs_tnc_sigs = fread("../../../Data/External/PCAWG/PCAWG_sigProfiler_SBS_signatures_in_samples.csv")

sbs_tnc_sigs_mat = as.matrix(sbs_tnc_sigs[, 4:ncol(sbs_tnc_sigs)])
sbs_tnc_sigs_mat_rel = sweep(sbs_tnc_sigs_mat,1, apply(sbs_tnc_sigs_mat, 1, sum), "/")
rownames(sbs_tnc_sigs_mat_rel) = sbs_tnc_sigs$`Sample Names`
```

## Trinucleotide signature contributions estimated by deconstructSigs on PCAWG trinucletide counts

```{r}
# These are SBS contributions estimated by deconstructSigs on the same samples using TNC counts from PCAWG
sbs_tnc_sigs_with_LeeSix_env = new.env()
load("../../../Data/PCAWG/Signature_weights_PCAWG_WGS_cohort_Alexandrov2018_and_LeeSix.Rdata", sbs_tnc_sigs_with_LeeSix_env)

sbs_tnc_sigs_deconstructSigs_mat = get("signature_weight_mat", sbs_tnc_sigs_with_LeeSix_env)
rownames(sbs_tnc_sigs_deconstructSigs_mat) = ed[rownames(sbs_tnc_sigs_deconstructSigs_mat), "specimenID"]
```

## POLE/D1 mutations from the PCAWG project (ICGC)

```{r}
pole_d_variants = fread("../../../Data/External/PCAWG/POLE_D_variants_ICGC.txt")
pole_d_variant_samples = unique(pole_d_variants$sampleID)

sample_env = new.env()
load("../../../Data/External/PCAWG/ICGC_DCC_release26_projects_donors_samples_specimen.Rdata", envir=sample_env)

sample_desc = get("samples", sample_env)
setkey(sample_desc, icgc_sample_id)

pole_d_variant_specimen = unique(sample_desc[pole_d_variant_samples]$icgc_specimen_id)
ed$POLE.D_mutant = ifelse(ed$specimenID %in% pole_d_variant_specimen, "POLE/D1 mutated", "POLE/D1 WT")
```

```{r}
ed_specimen = ed
rownames(ed_specimen) = ed_specimen$specimenID
```

# Data overview

## SBS trinucleotide contributions across all samples and cancer entities

```{r, fig.width=16, fig.height=12}
pheatmap(sbs_tnc_sigs_mat_rel, annotation_row = ed_specimen, cluster_cols = F, show_rownames = F)
```

# Projection of trinucleotide signatures to pentanucleotide change space

For each signature of interest we select the trinucleotide changes that contribute strongest to the signature.

```{r}
source("../shared_scripts/trinuc_signaturesl.R")

tmp = unlist(alexandrov_and_lee_six_sigs["SBS28",])
tmp = tmp[order(-tmp)]
tmp_cs = cumsum(tmp)
SBS28_tc = names(tmp[tmp>0.02 & tmp_cs<0.8])

SBS28_tc_weights = tmp[SBS28_tc]

tmp = unlist(alexandrov_and_lee_six_sigs["SBS41",])
tmp = tmp[order(-tmp)]
tmp_cs = cumsum(tmp)
SBS41_tc = names(tmp[tmp>0.02 & tmp_cs<0.8])
SBS41_tc_weights = tmp[SBS41_tc]

tmp = unlist(alexandrov_and_lee_six_sigs["SBS17b",])
tmp = tmp[order(-tmp)]
tmp_cs = cumsum(tmp)
SBS17b_tc = names(tmp[tmp>0.02 & tmp_cs<0.9])
SBS17b_tc_weights = tmp[SBS17b_tc]

tmp = unlist(alexandrov_and_lee_six_sigs["SBS26",])
tmp = tmp[order(-tmp)]
tmp_cs = cumsum(tmp)
SBS26_tc = names(tmp[tmp>0.02 & tmp_cs<0.9])
SBS26_tc_weights = tmp[SBS26_tc]

tmp = expected_mut_CAM[order(-expected_mut_CAM$CAM6),]
tmp$cs = cumsum(tmp$CAM6)
CAM6_tc = as.character(subset(tmp, CAM6>0.00 & cs<0.8)$tc)
```



```{r, fig.width=12, fig.height=6}
xx = apply(alexandrov_and_lee_six_sigs[, CAM6_tc], 1, function(x) cor(x, expected_mut_CAM[CAM6_tc,"CAM6"], method="spearman", use="complete") )
barplot(sort(xx, decreasing = T)[1:15], las=2, main="Top 15 signatures with highest overlap at TNC", ylab="Spearman correlation")
```

## CAM6 pentanucleotide change signature

```{r}
CAM6_pnc_tab_text = "CAM_motif,PN,weight
AAATTT,AATTT,2
AAATTT,ATTTA,0.5
AAATTT,ATTTC,0.5
AAATTT,ATTTG,0.5
AAATTT,ATTTT,0.5
AAATTT,ATTAA,0.125
AAATTT,ATTAC,0.125
AAATTT,ATTAG,0.125
AAATTT,ATTAT,0.125
AAATTT,ATTCA,0.125
AAATTT,ATTCC,0.125
AAATTT,ATTCG,0.125
AAATTT,ATTCT,0.125
AAATTT,ATTGA,0.125
AAATTT,ATTGC,0.125
AAATTT,ATTGG,0.125
AAATTT,ATTGT,0.125
AAATTT,ATTTA,0.125
AAATTT,ATTTC,0.125
AAATTT,ATTTG,0.125
AAATTT,ATTTT,0.125
AAAATT,AATTT,1
AAAATT,AATTA,0.25
AAAATT,AATTC,0.25
AAAATT,AATTG,0.25
AAAATT,AATTT,0.25
AAAATT,ATTAA,0.0625
AAAATT,ATTAC,0.0625
AAAATT,ATTAG,0.0625
AAAATT,ATTAT,0.0625
AAAATT,ATTCA,0.0625
AAAATT,ATTCC,0.0625
AAAATT,ATTCG,0.0625
AAAATT,ATTCT,0.0625
AAAATT,ATTGA,0.0625
AAAATT,ATTGC,0.0625
AAAATT,ATTGG,0.0625
AAAATT,ATTGT,0.0625
AAAATT,ATTTA,0.0625
AAAATT,ATTTC,0.0625
AAAATT,ATTTG,0.0625
AAAATT,ATTTT,0.0625
AAAATT,ATTTT,1
AAAATT,TTTTA,0.25
AAAATT,TTTTC,0.25
AAAATT,TTTTG,0.25
AAAATT,TTTTT,0.25
AAAATT,TTTAA,0.0625
AAAATT,TTTAC,0.0625
AAAATT,TTTAG,0.0625
AAAATT,TTTAT,0.0625
AAAATT,TTTCA,0.0625
AAAATT,TTTCC,0.0625
AAAATT,TTTCG,0.0625
AAAATT,TTTCT,0.0625
AAAATT,TTTGA,0.0625
AAAATT,TTTGC,0.0625
AAAATT,TTTGG,0.0625
AAAATT,TTTGT,0.0625
AAAATT,TTTTA,0.0625
AAAATT,TTTTC,0.0625
AAAATT,TTTTG,0.0625
AAAATT,TTTTT,0.0625
AATATT,AATAT,2
AATATT,TATTA,0.5
AATATT,TATTC,0.5
AATATT,TATTG,0.5
AATATT,TATTT,0.5
AATATT,ATTAA,0.125
AATATT,ATTAC,0.125
AATATT,ATTAG,0.125
AATATT,ATTAT,0.125
AATATT,ATTCA,0.125
AATATT,ATTCC,0.125
AATATT,ATTCG,0.125
AATATT,ATTCT,0.125
AATATT,ATTGA,0.125
AATATT,ATTGC,0.125
AATATT,ATTGG,0.125
AATATT,ATTGT,0.125
AATATT,ATTTA,0.125
AATATT,ATTTC,0.125
AATATT,ATTTG,0.125
AATATT,ATTTT,0.125
"

CAM6_pnc_tab = as.data.table(read.table(text=CAM6_pnc_tab_text, sep=",", header=T, stringsAsFactors = F))
CAM6_pnc_tab$weight_overall = CAM6_pnc_tab$weight/sum(CAM6_pnc_tab$weight)

CAM6_pnc_final = CAM6_pnc_tab[, .(CAM6_weight=sum(weight_overall)), by="PN"]
```

## Pentanucleotide annotation table

```{r}
pnc_anno = data.frame(pn=pnc_data$Pentanucleotide, pnc=paste0(substr(pnc_data$Pentanucleotide,1,2),"[",pnc_data$`Mutation type`,"]", substr(pnc_data$Pentanucleotide,4,5)), stringsAsFactors = F)
pnc_anno$tnc = substr(pnc_anno$pnc, 2,8)
pnc_anno$nc = substr(pnc_anno$pnc,4,6)

pnc_anno$tn = substr(pnc_anno$pn, 2,4)

pnc_anno = merge(pnc_anno, CAM6_pnc_final, by.x="pn", by.y="PN", all.x=T, sort=F)
pnc_anno$CAM6_weight = ifelse(is.na(pnc_anno$CAM6_weight), 0, pnc_anno$CAM6_weight)

pnc_anno$SBS28_weights_raw=ifelse(pnc_anno$tnc %in% names(SBS28_tc_weights), SBS28_tc_weights[pnc_anno$tnc], 0)
pnc_anno$SBS41_weights_raw=ifelse(pnc_anno$tnc %in% names(SBS41_tc_weights), SBS41_tc_weights[pnc_anno$tnc], 0)
pnc_anno$SBS17b_weights_raw=ifelse(pnc_anno$tnc %in% names(SBS17b_tc_weights), SBS17b_tc_weights[pnc_anno$tnc], 0)
pnc_anno$SBS26_weights_raw=ifelse(pnc_anno$tnc %in% names(SBS26_tc_weights), SBS26_tc_weights[pnc_anno$tnc], 0)

pnc_anno$SBS28_weights = pnc_anno$SBS28_weights_raw/sum(pnc_anno$SBS28_weights_raw)
pnc_anno$SBS41_weights = pnc_anno$SBS41_weights_raw/sum(pnc_anno$SBS41_weights_raw)
pnc_anno$SBS17b_weights = pnc_anno$SBS17b_weights_raw/sum(pnc_anno$SBS17b_weights_raw)
pnc_anno$SBS26_weights = pnc_anno$SBS26_weights_raw/sum(pnc_anno$SBS26_weights_raw)

rownames(pnc_anno) = pnc_anno$pnc
```

# Correlation of pentanucleotide change contributions to CAM6 signatures and to SBS trinucleotide contributions

```{r}
cam6_w_ordered = pnc_anno[colnames(pnc_mat_rel),]$CAM6_weight
```

# Heatmaps

```{r}
tmp = ed
tmp$fullID = rownames(ed)
tmp = merge(tmp, sbs_tnc_sigs_mat_rel[,c("SBS28","SBS41","SBS57","SBS26","SBS40","SBS17b"), drop=F], by.x="specimenID",by.y=0, all.x=T, sort=F)
tmp = merge(tmp, sbs_tnc_sigs_deconstructSigs_mat[, c("N2"),drop=F], by.x="specimenID", by.y=0, all.x=T, sort=F)
rownames(tmp) = tmp$fullID
tmp$SBS28_detection = ifelse(tmp$SBS28>0, "SBS28","nd")
tmp$SBS41_detection = ifelse(tmp$SBS41>0, "SBS41","nd")
tmp$SBS57_detection = ifelse(tmp$SBS57>0, "SBS57","nd")
tmp$SBS17b_detection = ifelse(tmp$SBS17b>0, "SBS17b","nd")
tmp$SBS26_detection = ifelse(tmp$SBS26>0, "SBS26","nd")
tmp$SBS40_detection = ifelse(tmp$SBS40>0, "SBS40","nd")
tmp$SBSA_detection = ifelse(tmp$N2>0.05, "SBSA","nd")

zz = apply(pnc_mat_rel, 1, function(x) sum(x*cam6_w_ordered))
tmp$CAM6_weighted_sum = 0
tmp[names(zz),]$CAM6_weighted_sum = zz
tmp$COREAD = ifelse(grepl("ColoRect-AdenoCA", rownames(tmp)), "yes","no")
tmp$UCEC = ifelse(grepl("Uterus-AdenoCA", rownames(tmp)), "yes","no")
row_anno = tmp[, c("POLE.D_mutant","cancer_entity","SBS28_detection", "SBS40_detection","SBS26_detection","SBS17b_detection", "CAM6_weighted_sum","SBSA_detection", "COREAD","UCEC")]

col_anno = pnc_anno[, c("pn","SBS28_weights", "SBS41_weights", "CAM6_weight", "SBS17b_weights", "SBS26_weights")]
col_anno$s5  = substr(col_anno$pn, 1,2)
col_anno$s3  = substr(col_anno$pn, 4,5)


anno_colors = list(POLE.D_mutant=c("POLE/D1 WT"="white","POLE/D1 mutated"="black"), "SBS28_detection"=c("SBS28"="red","nd"="white"), "SBS41_detection"=c("SBS41"="red","nd"="white"),"SBS40_detection"=c("SBS40"="red","nd"="white"), "SBS26_detection"=c("SBS26"="red","nd"="white"), "SBS17b_detection"=c("SBS17b"="red","nd"="white"),"SBSA_detection"=c("SBSA"="green","nd"="white"), "COREAD"=c("yes"="blue","no"="white"), "UCEC"=c("yes"="brown","no"="white"))

```

```{r}
xx = subset(pnc_anno, CAM6_weight>0.02| SBS28_weights  > 0 | SBS17b_weights > 0)
xx = xx[order(xx$pn,xx$nc),]

tmp_mat = pnc_mat_rel[, xx$pnc]
```

```{r, fig.width=18, fig.height=10}
row_anno2 = row_anno[, c("POLE.D_mutant","SBS28_detection","SBS17b_detection","SBSA_detection","CAM6_weighted_sum")]
col_anno2 = pnc_anno[, c("SBS28_weights","SBS17b_weights", "CAM6_weight")]
colnames(col_anno2) = gsub("CAM6_weight","AAWWTT weights",colnames(col_anno2))
colnames(row_anno2) = gsub("CAM6_weighted_sum","AAWWTT PN signature",colnames(row_anno2))
tmp_mat_sel = tmp_mat[rownames(subset(ed, cancer_entity=="ColoRect-AdenoCA")),]
```

```{r, fig.width=18, fig.height=10}
tmp_mat_sel_ordered = tmp_mat_sel[, order(-pnc_anno[colnames(tmp_mat_sel),"SBS28_weights"],-pnc_anno[colnames(tmp_mat_sel),"CAM6_weight"])]
rr = pheatmap(tmp_mat_sel_ordered, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = F, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", main="Colorectal cancer only, SBS28 PNC", fontsize_col = 8, gaps_col = sum(pnc_anno$SBS28_weights>0))
```


```{r, fig.width=18, fig.height=10}
dev.off()
pdf(file="../../../Results/4_CDM-association-with-cancer/SBS28_in_COAD_WGS_PCAWG.pdf", width=18,height=10)
#heatmap(tmp_mat_sel, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = T, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", fontsize_col = 8)
pheatmap(tmp_mat_sel_ordered, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = F, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", main="", fontsize_col = 8, gaps_col = sum(pnc_anno$SBS28_weights>0))
dev.off()
```


```{r, fig.width=16, fig.height=10}
tmp_mat_sel = tmp_mat[rownames(subset(ed, cancer_entity=="Stomach-AdenoCA")),]
tmp_mat_sel = tmp_mat_sel[apply(tmp_mat_sel,1,sum)>0,]
```


```{r, fig.width=16, fig.height=10}
tmp_mat_sel_ordered = tmp_mat_sel[, order(-pnc_anno[colnames(tmp_mat_sel),"SBS17b_weights"],-pnc_anno[colnames(tmp_mat_sel),"CAM6_weight"])]
rr = pheatmap(tmp_mat_sel_ordered, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = F, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", main="Stomach-AdenoCA only", fontsize_col = 8, gaps_col = sum(pnc_anno$SBS17b_weights>0))
```

```{r, fig.width=18, fig.height=10}
dev.off()
pdf(file="../../../Results/4_CDM-association-with-cancer/SBS17_in_STAD_WGS_PCAWG.pdf", width=18,height=10)
#pheatmap(tmp_mat_sel, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = T, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", main="Stomach-AdenoCA only", fontsize_col = 8)
pheatmap(tmp_mat_sel_ordered, annotation_row = row_anno2, annotation_col = col_anno2, show_rownames = F, cluster_cols = F, cluster_rows=T, annotation_colors = anno_colors, clustering_distance_rows = "correlation", main="", fontsize_col = 8, gaps_col = sum(pnc_anno$SBS17b_weights>0))
dev.off()
```

# Software versions

```{r}
sessionInfo()
```

