---
title: "Nucleotide enrichment in windows DSB+/-7nt  - pentanucleotide patterns"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: show
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, message=FALSE}
rm(list=ls())
library(Biostrings)
library(ggplot2)
library(data.table)
```

# Introduction

Nucleotide enrichment data for pentanucleotides for all replicates and conditions: E.coli WT infection, E.coli mutant infection and controls are imported.  

Since we noticed that many of the top ranking hits are similar to barcodes used in BLISS for those samples we mark nucleotide patterns which match known barcode patterns in forward or reverse complement direction with at most one mismatch as Barcode associated. Those patterns will be excluded.

# Importing data and analysis
```{r}
pattern_length=5
input_folder = "../../Data/Sequence_enrichment/Pentanucleotides/"
```

```{r, message=FALSE, warning=FALSE}
source("./read_tables_and_barcodes.R")

bc_info = get_barcode_matches(pattern_length, "/data_genome2/projects/MB222-GuideSeq/BLISS/MPIIB_Pipeline/metadata/Barcodes_all.txt", max_mismatch=0)

all_nuc_enrichment_tabs = read_enrichment_tables(input_folder, bc_info)
```

```{r}
fill_manual = scale_fill_manual(values=c("AT4"="red3", "AT5"="royalblue3", "other" = "grey70"))
col_manual =  scale_color_manual(breaks=c("AT4", "AT5", "other"), values=c("red3", "royalblue3", "grey70"))
```

```{r}
all_nuc_enrichment_tabs_agg = all_nuc_enrichment_tabs[, .(Mean=mean(logRatio, na.rm=T), SD=sd(logRatio, na.rm=T), rep_cnt=sum(!is.na(logRatio)), bc=ifelse(any(barcode_match_library!="no BC"),"BC f/r", "no BC"), SEM_err_prop= sqrt(sum(sd_log2ratio**2)/sum(!is.na(logRatio))) ), by=c("Nt","condition", "at_cnt")]
all_nuc_enrichment_tabs_agg[, SEM:=pmax(SEM_err_prop, SD/sqrt(rep_cnt), na.rm=T ) ]
all_nuc_enrichment_tabs_agg[, CI:=SEM * qt(0.975, rep_cnt-1)]
setorder(all_nuc_enrichment_tabs_agg, condition, -Mean)

```

# Visualization

## WT vs Mutant
```{r}
sel_tab = subset(all_nuc_enrichment_tabs, condition == "WTvsMut")

sel_tab[, mean_logratio:=mean(logRatio), by="Nt"]

sel_tab_agg = subset(all_nuc_enrichment_tabs_agg, condition == "WTvsMut")
setorder(sel_tab_agg, -Mean)
sel_tab_agg[, Nt:=factor(Nt, levels=Nt)]

```

### Overview

Log2Ratio for each pattern (patterns similar to BC removed)

```{r, fig.width=8, fig.height=6}
tmp= sel_tab_agg[sel_tab_agg$bc=="no BC"]
tmp[, AT_content := ifelse(at_cnt==5, "AT5", ifelse(at_cnt==4, "AT4", "other"))]

p = ggplot(tmp, aes(x=Nt, y=Mean, fill=AT_content)) + geom_bar(stat="identity") +
  theme(axis.text.x = element_blank(), axis.ticks.x=element_blank(), text=element_text(size=28), axis.text.y=element_text(size=24), legend.position="none") +
  geom_hline(yintercept = 0, size=0.5) + coord_cartesian(ylim=c(-.4, 1.2), xlim=c(-2,nrow(tmp)+2)) + 
  fill_manual + labs(x="All Pentanucleotide sequences", y="Mean Log2 Ratio",  title="", subtitle = "") 


print(p)

```


```{r}
pdf("../../Results/2_DSB-sequence-specificity/PN_WTvsMut_all_bars.pdf", width=10, height = 6)
print(p)
dev.off()

```


### Top 50 patterns

Log2Ratio for top50 patterns (patterns similar to BC removed)

```{r, fig.width=14.5, fig.height=9}
tmp= sel_tab_agg[sel_tab_agg$bc=="no BC"  ] 
tmp[, AT_content := ifelse(at_cnt==5, "AT5", ifelse(at_cnt==4, "AT4", "other"))]
tmp = tmp[1:50]

tmp2 = sel_tab[as.character(sel_tab$Nt) %in% as.character(tmp$Nt), ]

p = ggplot(tmp, aes(x=Nt, y=Mean, fill=AT_content)) + geom_bar(stat="identity") + theme_classic() +
  geom_errorbar(width=.1, aes(ymin=Mean-CI, ymax=Mean+CI), colour="black", alpha=0.6) + 
  geom_point(data=tmp2, aes(x=Nt, y=logRatio), inherit.aes = F, size=1.5) + 
  theme(text=element_text(size=24), axis.text.x = element_text(angle = 60, hjust = 1, size=16), axis.text.y = element_text(size=24), legend.position="none") +
  fill_manual + coord_cartesian(ylim=c(-.5, 2)) + labs(x="Top 50 Pentanucleotide sequences", y="Mean Log2 Ratio", title="", subtitle = "") 
print(p)

```

```{r}
pdf("../../Results/2_DSB-sequence-specificity/PN_WTvsMut_top50_bars.pdf", width=14.5, height = 9)
print(p)
dev.off()
```


## Mut vs NT

```{r}
sel_tab = subset(all_nuc_enrichment_tabs, condition == "MutvsCtrl")

sel_tab[, mean_logratio:=mean(logRatio), by="Nt"]

sel_tab_agg = subset(all_nuc_enrichment_tabs_agg, condition == "MutvsCtrl")
setorder(sel_tab_agg, -Mean)
sel_tab_agg[, Nt:=factor(Nt, levels=Nt)]

```

### Overview

Log2Ratio for each pattern (patterns similar to BC removed)

```{r, fig.width=10, fig.height=6}
tmp= sel_tab_agg[sel_tab_agg$bc=="no BC"]
tmp[, AT_content := ifelse(at_cnt==5, "AT5", ifelse(at_cnt==4, "AT4", "other"))]

p = ggplot(tmp, aes(x=Nt, y=Mean, fill=AT_content)) + geom_bar(stat="identity") +
  theme(axis.text.x = element_blank(), axis.ticks.x=element_blank(), text=element_text(size=28), axis.text.y=element_text(size=24), legend.position="none") +
  geom_hline(yintercept = 0, size=0.5) + coord_cartesian(ylim=c(-.4, 1.2), xlim=c(-1,nrow(tmp)+1)) + 
  fill_manual + labs(x="All Pentanucleotide sequences", y="Mean Log2 Ratio",  title="", subtitle = "") 

print(p)

```

```{r}
pdf("../../Results/2_DSB-sequence-specificity/PN_MutvsNT_all_bars.pdf", width=10, height = 6)
print(p)
dev.off()

```


### Top 50 patterns

Log2Ratio for top50 patterns (patterns similar to BC removed)

```{r, fig.width=14.5, fig.height=9}
tmp= sel_tab_agg[sel_tab_agg$bc=="no BC"  ]
tmp[, AT_content := ifelse(at_cnt==5, "AT5", ifelse(at_cnt==4, "AT4", "other"))]
tmp = tmp[1:50]

tmp2 = sel_tab[as.character(sel_tab$Nt) %in% as.character(tmp$Nt), ]

p = ggplot(tmp, aes(x=Nt, y=Mean, fill=AT_content)) + geom_bar(stat="identity") + theme_classic() +
  geom_errorbar(width=.1, aes(ymin=Mean-CI, ymax=Mean+CI), colour="black", alpha=0.6) + 
  geom_point(data=tmp2, aes(x=Nt, y=logRatio), inherit.aes = F, size=1.5) + 
  theme(text=element_text(size=24), axis.text.x = element_text(angle = 60, hjust = 1, size=16), axis.text.y = element_text(size=24), legend.position="none") +
  fill_manual + coord_cartesian(ylim=c(-.5, 2)) + labs(x="Top 50 Pentanucleotide sequences", y="Mean Log2 Ratio", title="", subtitle = "")

print(p)

```

```{r}
pdf("/data_genome2/projects/MB222-GuideSeq/GitHub_2020/Results/2_DSB-sequence-specificity/PN_MutvsNT_top50_bars.pdf", width=14.5, height = 9)
print(p)
dev.off()
```


# Software versions

```{r}
sessionInfo()
```

