---
title: "Finnish WGS cohort analysis - Alexandrov 2018 (COSMICv3) and Lee-Six et al. signatures"
author: "Hilmar Berger"
output:
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, message=FALSE}
rm(list=ls())
library(data.table)
library(Biostrings)
library(pheatmap)

library(deconstructSigs)
library(BiocParallel)

library(ggplot2)
library(naturalsort)
```

# Overview

We want to test if pentanucleotides are somehow associated to known or novel mutational signatures.
We will use the trinucleotide change frequencies as provided by Riku Katainen and Lauri Aaltonen.

```{r, message=FALSE}
data = fread("../../../Data/WGS_cohort/SBS_Trinucleotide_signature_contributions/Trinucleotide_mutcounts_CRC_WGS_n200.txt", sep="\t")
trinuc_change_mat = as.matrix(data[, 2:ncol(data)])
tmp = data$`#mutType`

# the notation of trinucleotide changes does not follow the general standard from e.g. COSMIC signatures. Fix notation. 
nuc_change = paste0(substr(tmp,2,2),">",substr(tmp,6,6))

plus_strand_nuc_changes_accepted = c("C>A","C>G","C>T","T>A","T>C","T>G")

revcomp_nuc_change = paste0(as.character(complement(DNAStringSet(substr(tmp,2,2)))),">",as.character(complement(DNAStringSet(substr(tmp,6,6)))))
revcomp_trinuc = as.character(reverseComplement(DNAStringSet(substr(tmp, 1, 3))))

trinuc_change = ifelse(nuc_change %in% plus_strand_nuc_changes_accepted, 
                       paste0(substr(tmp, 1,1),"[",nuc_change,"]", substr(tmp,3,3)), 
                       paste0(substr(revcomp_trinuc, 1,1),"[",revcomp_nuc_change,"]", substr(revcomp_trinuc,3,3)) )

rownames(trinuc_change_mat) = trinuc_change
trinuc_change_mat = trinuc_change_mat[order(rownames(trinuc_change_mat)),]

ed = data.frame(sampleID=colnames(data)[2:ncol(data)], stringsAsFactors = F)
tt = regmatches(ed$sampleID, regexpr("(.+)_T_", ed$sampleID))
ed$ID = make.names(substr(tt, 1, nchar(tt)-3))
tt2 = sapply(strsplit(ed$sampleID,"-"), `[`,1 )
ed$group = ifelse(tt2 %in% c("POLE","MSI"), tt2, "other" )
rownames(ed) = ed$sampleID
```

# Known signatures

We use package deconstructSigs to estimate the contribution of the mutational signatures published by Alexandrov 2018 in all samples. We add new signatures derived from the Lee-Six/Stratton paper on signatures in colorectal cells.

```{r}
new_sigs = read.csv(file="../../../Data/External/SBS_signatures/sigProfiler_SBS_signatures_2018_03_28.csv")
new_sigs_fixed = as.data.frame(t(new_sigs[, 3:ncol(new_sigs)]))
colnames(new_sigs_fixed) = apply(new_sigs[, 1:2], 1, function(x) {paste0(substr(x[2],1,1),"[", x[1],"]", substr(x[2],3,3))})
```

## Stratton colon epithelium signature

We additionally add signatures extracted from the supplementary data (i.e. git repository) of the Lee-Six paper. There we used the provided code for using a Hierarchical Dirichlet Process (HDP) method to extract SBS signatures.

```{r}
lee_six_sigs = read.table("../../../Data/External/SBS_signatures/Signature_category_counts_LeeSix.txt", sep="\t", header=T)

lee_six_mat = as.matrix(lee_six_sigs[, 2:ncol(lee_six_sigs)])
tmp_c = as.character(lee_six_sigs$X)
nuc_change = gsub("\\.",">", substr(tmp_c, 1, 3))
trinuc_change = paste0(substr(tmp_c,8,8),"[",nuc_change,"]",substr(tmp_c,10,10))

rownames(lee_six_mat) = trinuc_change

lee_six_mat_new_sigs = lee_six_mat[, c(paste0("N",1:5))]

alexandrov_and_lee_six_sigs = as.matrix(rbind(t(lee_six_mat_new_sigs), new_sigs_fixed[,rownames(lee_six_mat_new_sigs)]))

```

## A-priori similarity of all signatures

```{r, fig.width=10, fig.height=10}
cosineDist <- function(x){
  as.dist(1 - x%*%t(x)/(sqrt(rowSums(x^2) %*% t(rowSums(x^2))))) 
}

tmp2 = as.matrix(cosineDist(alexandrov_and_lee_six_sigs))
pheatmap(1-tmp2, main="Signature similarity (cosine correlation)")
```


```{r}
set.seed(19825)
reprocess_data = T
if(reprocess_data) {

  sigs.input = as.data.frame(t(trinuc_change_mat))
  ed_sigs_input = ed
  
  all_donors = sort(rownames(sigs.input))
  all_donors = all_donors[(apply(sigs.input, 1, sum)>0)]
  
  sigs.input_selected = sigs.input[all_donors,]
  
  sig_fun = function(sample_id) {
    sigs_found = NA
    try((sigs_found = whichSignatures(tumor.ref = sigs.input_selected, 
                     signatures.ref = as.data.frame(alexandrov_and_lee_six_sigs), 
                     sample.id = sample_id, contexts.needed=TRUE, tri.counts.method = 'default', signature.cutoff = 0.02)) )
    return(sigs_found)
  }
  
  all_sigs = bplapply(all_donors, sig_fun, BPPARAM=MulticoreParam(workers=24) )

  save(sigs.input, all_sigs, ed_sigs_input, file="../../../Data/WGS_cohort/SBS_Trinucleotide_signature_contributions/All_signatures_deconstructSigs_global_Alexandrov_LeeSix.Rdata")
} else {
  load("../../../Data/WGS_cohort/SBS_Trinucleotide_signature_contributions/All_signatures_deconstructSigs_global_Alexandrov_LeeSix.Rdata")
}
```


```{r}
all_donors = sort(rownames(sigs.input))
all_donors = all_donors[(apply(sigs.input, 1, sum)>0)]

failed_samples = (1:length(all_sigs))[!unlist(lapply(all_sigs, is.list))]

all_sigs_filtered = all_sigs
all_donors_final = all_donors
if(length(failed_samples)>0) {
  all_sigs_filtered[failed_samples] <- NULL
  all_donors_final = all_donors[-failed_samples]
}

names(all_sigs_filtered) = all_donors_final

signature_weight_mat = as.matrix(do.call(rbind, lapply(all_sigs_filtered, function(x) x$weights)))
```

# Signature contributions across all samples and contexts

```{r, fig.width=16, fig.height = 10}

rownames(ed_sigs_input) = ed_sigs_input$TCGA_donor

col_anno = ed_sigs_input[,c("group"),drop=F]
rownames(col_anno) = ed_sigs_input$sampleID

pheatmap(t(signature_weight_mat), cluster_cols=T, show_colnames = F, annotation_col = col_anno)

```


```{r}
sig_mat_ts = reshape2::melt(signature_weight_mat)
colnames(sig_mat_ts) = c("ID", "Signature", "Contribution")
sig_mat_ts$sampleID = sig_mat_ts$ID
sig_mat_ts$cancer_entity = ed[as.character(sig_mat_ts$sampleID), "group"]
```

```{r, fig.width=15, fig.height=12}
ggplot(sig_mat_ts, aes(x=Signature, y=Contribution)) + geom_boxplot(outlier.shape = 20, outlier.alpha = 0.3) + facet_grid(cancer_entity~.) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


# Correlation of signature contributions across samples

```{r, fig.height=8}
cc = cor(signature_weight_mat[,apply(signature_weight_mat, 2, sum)>0], method="spearman")
pheatmap(cc, main="Correlation of signature contribution across samples")
```


```{r}
save(signature_weight_mat, file="../../../Data/WGS_cohort/SBS_Trinucleotide_signature_contributions/Signature_weights_WGS_cohort_Alexandrov2018_and_LeeSix.Rdata")
```

