---
title: "Theoretical SBS signature"
author: "Hilmar Berger"
date: "3/19/2020"
output: html_document
---

```{r}
rm(list=ls())
library(ggplot2)
library(ggthemes)
library(pheatmap)
source("../shared_scripts/trinuc_signaturesl.R")
```

# Overview

We here use a theoretically derived SBS trinucleotide change signature based on trinucleotide frequencies in the colibactin associated AAWWTT motif combined with assumptions on a uniform distribution of T>[C/G/A] nucleotides. 

A further extension is the inclusion of hotspots at positions 2/5 of the motif and a preference for A:T>G:C mutations at those positions. 

# CAM6 SBS signature with uniform distribution

```{r}

e_uniform = expected_mut_CAM[, c("tc","nc","CAM6")]
colnames(e_uniform) = c("tc","nc","weight")
e_uniform$signature = "Theoretical AAWWTT signature (uniform distribution)"

sbsa = data.frame(tc=rownames(lee_six_mat_new_sigs), nc=substr(rownames(lee_six_mat_new_sigs), 3,5), weight= lee_six_mat_new_sigs[, "N2"], signature="SBSA")

e_uniform_and_n2 = rbind(e_uniform, sbsa)
e_uniform_and_n2$tc = factor(e_uniform_and_n2$tc, levels=trinuc_changes_ordered$tc)
e_uniform_and_n2$signature = factor(e_uniform_and_n2$signature, levels=c("Theoretical AAWWTT signature (uniform distribution)","SBSA"))
```

```{r, fig.width=10, fig.height=4}
p = ggplot(e_uniform_and_n2, aes(x=tc, y=weight, fill=nc)) + geom_bar(stat="identity") + facet_wrap(~signature, ncol=1) + theme_few() + theme(text=element_text(size=20), axis.text.x=element_text(angle=90, hjust = 0, size=5)) + labs(x="", fill="Nucleotide\nchange", y = "Proportion") 
print(p)
```

```{r}
pdf("../../../Results/4_CDM-association-with-cancer/Theoretical_AAWWTT_SBS_signature.pdf", width = 10, height = 4)
print(p)
dev.off()
```

```{r}
cosine_func <- function(x,y) x %*% y / sqrt(x%*%x * y%*%y)
cc1 = apply(alexandrov_and_lee_six_sigs,1, cosine_func, expected_mut_CAM$CAM6)
cc1_res = as.data.frame(cc1)
cc1_res$signature=rownames(cc1_res)
colnames(cc1_res) = c("Cosine correlation")
head(cc1_res[order(-cc1_res$`Cosine correlation`),,drop=F],20)
```

```{r, fig.height=6}
tmp = cbind(cc1,cc1)
tmp = tmp[order(-tmp[, 1]),]
pheatmap(tmp[1:10,], cluster_rows = F, cluster_cols=F, cellwidth = 40, cellheight = 40, gaps_col = 1, fontsize_row = 20, fontsize = 14, show_colnames = F)
```


```{r, fig.width=6, fig.height=6}
dev.off()
cairo_pdf(file="../../../Results/4_CDM-association-with-cancer/Correlation_CAM6_SBSA_heatmap.pdf", width=6, height = 6)
pheatmap(tmp[1:10,], cluster_rows = F, cluster_cols=F, cellwidth = 40, cellheight = 40, gaps_col = 1, fontsize_row = 20, fontsize = 14, show_colnames = F)
dev.off()
```

# CAM6 with positional bias towards positions 2 and 5


```{r}

e_uniform = expected_mut_CAM[, c("tc","nc","CAM6_hotspot")]
colnames(e_uniform) = c("tc","nc","weight")
e_uniform$signature = "Theoretical AAWWTT signature (uniform distribution with hotspot)"

sbsa = data.frame(tc=rownames(lee_six_mat_new_sigs), nc=substr(rownames(lee_six_mat_new_sigs), 3,5), weight= lee_six_mat_new_sigs[, "N2"], signature="SBSA")

e_uniform_and_n2 = rbind(e_uniform, sbsa)
e_uniform_and_n2$tc = factor(e_uniform_and_n2$tc, levels=trinuc_changes_ordered$tc)
e_uniform_and_n2$signature = factor(e_uniform_and_n2$signature, levels=c("Theoretical AAWWTT signature (uniform distribution with hotspot)","SBSA"))
```

```{r, fig.width=10, fig.height=4}
p = ggplot(e_uniform_and_n2, aes(x=tc, y=weight, fill=nc)) + geom_bar(stat="identity") + facet_wrap(~signature, ncol=1) + theme_few() + theme(text=element_text(size=20), axis.text.x=element_text(angle=90, hjust = 0, size=5)) + labs(x="", fill="Nucleotide\nchange", y = "Proportion") 
print(p)
```

<!-- ```{r} -->
<!-- pdf("Theoretical_AAWWTT_SBS_signature_hotspots.pdf", width = 10, height = 4) -->
<!-- print(p) -->
<!-- dev.off() -->
<!-- ``` -->

```{r}
cosine_func <- function(x,y) x %*% y / sqrt(x%*%x * y%*%y)
cc1 = apply(alexandrov_and_lee_six_sigs,1, cosine_func, expected_mut_CAM$CAM6_hotspot)
cc1_res = as.data.frame(cc1)
cc1_res$signature=rownames(cc1_res)
colnames(cc1_res) = c("Cosine correlation")
head(cc1_res[order(-cc1_res$`Cosine correlation`),,drop=F],20)
```

```{r, fig.height=6}
tmp = cbind(cc1,cc1)
tmp = tmp[order(-tmp[, 1]),]
pheatmap(tmp[1:10,], cluster_rows = F, cluster_cols=F, cellwidth = 40, cellheight = 40, gaps_col = 1, fontsize_row = 20, fontsize = 14, show_colnames = F)
```


<!-- ```{r} -->
<!-- pdf(file="Correlation_CAM6_hotspot_2.5_heatmap.pdf", width=6, height = 6) -->
<!-- pheatmap(tmp[1:10,], cluster_rows = F, cluster_cols=F, cellwidth = 40, cellheight = 40, gaps_col = 1, fontsize_row = 20, fontsize = 14, show_colnames = F) -->
<!-- pheatmap(tmp[1:10,], cluster_rows = F, cluster_cols=F, cellwidth = 40, cellheight = 40, gaps_col = 1, fontsize_row = 20, fontsize = 14, show_colnames = F) -->
<!-- dev.off() -->
<!-- ``` -->